---
title: "Exploring the GSF data"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    df_print: "paged"
---

```{r, message = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
```

```{r}
d <- read_csv("GudK_T1_and_T2_fixed.csv")
```

```{r}
d
```

## Overview

We want to know about correctness and speed of responses. The main predictor is the `condition` variable. Here are a few plots.

### Proportion correct by condition

```{r, fig.width = 9, fig.height = 6}
d %>%
  ggplot(aes(x = condition, fill = correct)) +
  geom_bar(position = position_dodge()) +
  facet_wrap(vars(subject)) ->
  fig_responses

fig_responses
```

Most subjects did quite well. But **EB12AY**, **ED21AD**, and **HA28TH** got nearly half their answers wrong. It doesn't look as though one condition was consistently more difficult than the other.

### RT distributions by condition

Considering only correct responses.

I always have to google how to show distributions overlaid on each other with ggplot, despite having done it enough times. It's the use of `position_identity()` that trips me up. This prevents the default positioning of bars stacked on top of each other instead of overlaid.

I've applied an arbitrary RT cutoff to stop the horizontal scale from getting too stretched out.

```{r, fig.width = 9, fig.height = 6}
d %>%
  filter(correct, RT < 5) %>%
  ggplot(aes(x = RT, fill = condition)) +
  geom_histogram(
    position = position_identity(),
    alpha = 0.4,
    binwidth = 0.1
  ) +
  facet_wrap(vars(subject)) +
  labs(x = "RT (s)") ->
  fig_RT

fig_RT
```

Most look like classic RT distributions, with an early peak and a long tail. A few subjects seem to be more uniformly taking their time. Most of these are the same ones that got nearly half their answers wrong.

### Mean RT by condition and correctness

The full distributions might be masking subtle on-average differences. So we can look at a plot of means as well.

Varying the vertical axis scale by subject (with `scales = "free_y"`) gives a closer view of each subject's pattern.

Since a plot of means is less cluttered than a full histogram, we can bring back the incorrect responses just in case those are of interest.

```{r, fig.width = 9, fig.height = 6, message = FALSE}
d %>%
  ggplot(aes(x = condition, y = RT, color = correct)) +
  stat_summary(
    position = position_dodge(width = 0.2),
    fun.data = mean_sdl,
    fun.args = c(mult = 1)
  ) +
  facet_wrap(vars(subject), scales = "free_y") +
  labs(y = "mean RT (s)", caption = "mean ±1 SD") ->
  fig_mean_RT

fig_mean_RT
```

Subjects tend to be faster with correct responses. Makes sense. There doesn't seem to be a consistent difference between conditions.

## Hand words

Curiously, there isn't actually a variable recording the correct word for each trial. This is my omission, something I forgot in the Python program for the experiment. But we can piece the words together from the gap text variable (e.g. *n_se*) and the correct letter variable (e.g. *o*).

```{r}
d %>%
  mutate(full_answer = str_replace(target, "_", corrAns)) ->
  d

d %>%
  select(full_answer) %>%
  distinct()
```

Now we can mark the hand-related words with a new variable. *Point* and *wave* make sense. *Crawl* feels a bit more dubious, since it involves the legs as well and is arguably not primarily hand-focused. It might be interesting to think about how to identify hand words more systematically.

```{r}
hand_words <- c(
  "crawl",
  "point",
  "wave"
)

d %>%
  mutate(hand_word = full_answer %in% hand_words) ->
  d
```

Now we can see the same figure of mean reaction times as above, but with the hand word variable in place of condition.

```{r, fig.width = 9, fig.height = 6}
fig_mean_RT_hands <- (fig_mean_RT %+% d) + aes(x = hand_word)

fig_mean_RT_hands
```

No striking pattern. But maybe there aren't enough hand words to go on.

## RT cutoff

The solution you linked on [StackOverflow](https://stackoverflow.com/questions/61601426/remove-outlier-rows-by-column-and-factor-in-r) looks good, and I think you are close to making it work for your case. But this sort of thing can be done much more clearly using the `dplyr` package. Also, using column numbers instead of names is always a bit risky.

The first tricky part of the problem is that we need a separate mean and standard deviation per subject. We can solve this with `group_by()` and then `summarize()`.

```{r}
d %>%
  filter(correct) %>%
  group_by(subject) %>%
  summarize(.groups = "drop",
    RT_mean = mean(RT),
    RT_SD = sd(RT)
  ) ->
  subject_summaries
```

Then use the new variables to calculate a cutoff per subject.

```{r}
RT_SD_multiplier <- 2.5

subject_summaries %>%
  mutate(RT_cutoff = RT_mean + RT_SD_multiplier * RT_SD) ->
  subject_summaries

subject_summaries
```

Let's check with a plot. We can add the cutoffs to the plot of RT distributions from earlier.

```{r, fig.width = 9, fig.height = 6}
fig_RT +
  geom_vline(
    mapping = aes(xintercept = RT_cutoff),
    data = subject_summaries,
    linetype = "dashed"
  )
```

Now we can join this information back into the main data frame.

```{r}
d %>%
  left_join(subject_summaries) ->
  d
```

And use it with `filter()` in subsequent plots or analyses.

```{r}
d %>%
  filter(RT < RT_cutoff)
```
